Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-14 ms.tex
retreival edits

\subsection{Retrieval}
\label{sec:retrieval}
We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b}.   Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}. The transmission forward model is coupled with the PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

Our nominal model includes a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the C/O, and a gray cloud-top pressure.  We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.  We put a uniform prior on the atmospheric metallicity of $0.01 - 30\times$ solar based on the upper limit from $\S$\,\ref{sec:interior}.

The retrieval results are shown in Figure\,\ref{fig:retrieval}.  The metallicity distribution spans the full range allowed by our priors, with some preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. These are not particularly precise constraints because the metallicity and cloud height are degenerate \citep{griffith14}.  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.  The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).

We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.  We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.

@@

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O. In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$. We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.  Assuming  solar C/O and $3\,\times$ solar metallicity (the median retrieved value for the chemical equilbrium), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.  Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.

>>>>

\subsection{Retrieval}
\label{sec:retrieval}
We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b}.   Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}. The transmission forward model is coupled with the PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

Our nominal model included a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the C/O, a gray cloud-top pressure, and the planet's 10-bar radius.  We fixed the T-P profile morphology but scaled the irradiation temperature to allow for the unknown albedo and heat transport efficiency.  We put a uniform prior on the atmospheric metallicity of $0.01 - 30\times$ solar based on the upper limit from $\S$\,\ref{sec:interior}.

The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).  There is one $3.4\sigma$ outlier data point (1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.  We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.  These models did not significantly improve the fit quality, and the main retrieval results were unchanged. We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.

The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.  The metallicity distribution spans the full range allowed by our priors, with some preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. These are not particularly precise constraints because the metallicity and cloud height are somewhat degenerate \citep{griffith14}.  The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.

@@

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.  Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a level we can detect.  In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$. We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.  Assuming  solar C/O and $3\,\times$ solar metallicity (the median value for the chemical equilbrium retrieval), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.  Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
